### 2020-04-15 intro.tex
Layout figs
fixed intro

Titan was quite an interesting machine from a reliability perspective.
Its operation was forced to execute three very significant rework cycles, two on
the mechanical assembly affecting the PCIe connector from the GPU
daughtercard to the motherboard, and a replacement of about 11,000 GPU
assemblies because of a failing resistor on a printed circuit board.

Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles. It shows the number of GPU changes at periodic inventories.
\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Daily volume of GPU swaps with labels on most frequent dates.}
  \label{fig:chronology}
\end{figure*}

We have conducted a good bit of statistical analysis on the failure
rates of these GPU assemblies, but I want to propose a different topic
that I have not seen for other/similar systems.  I want to complete
the statistical analysis of the failure rates as we near end of life,
looking for the opposite side of the “bathtub curve”.  There is
specific decay associated with the electronic components where, as
these products age, they will eventually fail thresholds for guard
band margin and other things. In a lot of cases, they will just fail
as well.

This machine is quite large, was very important to many people, and I
think this analysis could be very intriguing.

I am looking for help, very specifically with the statistics portion
of this, i.e. what do the failure rates for specific components look
like? What distributions do those failure rates follow? How could
appropriate models extrapolate failure rates beyond the end of service
date (the dotted lines) ?

I have submitted an abstract to a Cray-focused user group and workshop
on this topic, and it has been accepted.  I would like to execute this
statistical analysis and generate a short presentation (30 minutes)
and a short paper (4-8 pages is a likely target).  Presentation is
targeted for May 7-9 (TBD). An initial/working version of the paper
would be due earlier. If we miss that target (April 12), no big
deal. We can finish on a similar schedule as the presentation.  I have
all of the raw data for the component failures, and need the
statistical help to complete the analysis.

Because there is a huge swing in the failure rates of the GPUS from
2016/17 when the resistor problem escalated rapidly out of control, it
is statistically interesting, where we inject fresh material in to an
existing machine. We’d likely want to categorically separate new GPU
assemblies from old and could track the early life failures from those
new 11,000 parts versus the continued failure of the remaining 6000.

George, Christian-  any interest in this? Barney gave it a thumbs
up. I think this is a timely and very interesting topic.

\fix{Need a description of Titan GPU events timeline. Jim or Don?}

>>>>

From a realiability perspective, Titan was a very interesting
machine. Its operation was forced to execute three very significant
rework cycles, two on the mechanical assembly affecting the PCIe
connector from the GPU daughtercard to the motherboard, and a
replacement of about 11,000 GPU assemblies because of a failing
resistor on a printed circuit board. We write primarily about the epoch
that includes this last rework cycle. This includes Titan's most
stable and failure free period and contains the most reliable data on
GPU operation.

Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles and stable periods as indicated by the number of GPU changes at
periodic inventories.
\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology001.pdf}
  \caption{Daily volume of GPU swaps with labels on most frequent dates.}
  \label{fig:chronology}
\end{figure*}
After an initial fairly stable break-in period, the first rework cycle
begins in early 2013 and ends by May. \fix{Details?}. The second
rework cycle begins in late August of 2013 and ends by mid
December. This is followed by a very long and stable period until
about mid 2016, when GPU failures begin to rise. This results in the
final rework cycle, replacing 11,000 GPUs throughout all of 2017.

While the 2016 resistor failures were a surprise, there is a known
phenomenon \cite{referenceneeded} that describes a two phase process
that leads to failure. It also explains why there was a very stable
period preceeding the failures. There is specific decay associated
with the electronic components where, as these products age, they will
eventually fail thresholds for guard band margin and other things. In
a lot of cases, they will just fail as well. \fix{Give a reference and
  expand?}

Our analysis focuses on GPUs that were installed in the second rework
cycle and later, essentialy units installed near the beginning of 2014
and later.

We begin by describing related work in Sec.~\ref{sec:related}. Our
data collection is desctibed in Sec.~\ref{sec:dataprep} and the data
cleaning, checking, and lifetime summarization process in
Sec.~\ref{sec:dataclean}. A timme between failures analysis is given
in Sec.~\ref{sec:tbf}. Survival analysis based on Kaplan-Meier
modeling and Cox regression modeling of relative hazard rates are in
Sec.~\ref{sec:survival}. Finally, we discuss our main conclusions and
recommendations in Sec.~\ref{sec:discussion}.
